Make a README

### 2025-07-16 make_readme.qmd
Also state where deps can be obtained
Shorten README Chapter; add new in-depth page

Having settled on a license, it is time to add a final touch.
Imagine returning to your project in five years,
having forgotten most of the details of what you did exactly.
What would be useful to know in order to quickly understand
what is going on in the project?
This is what needs to be described in the README.
While you could just start writing along,
it is helpful to provide at least the following information in sections on their own.

### Name and Description

How is the project called?
What is it about?
Which files does the project folder contain?
How are they organized?

### Involved Data

Are any (empirical) data involved (e.g., being analyzed or used as input)?
From which sources can they be obtained?
Are they already included in the project folder?
Where is their data dictionary located?
Which terms, usage restrictions, or licenses apply?
If they are not publicly available,
is an alternative, synthesized version provided?

### Computational Requirements

What software needs to be installed to run the analysis
-- in other words, what are its dependencies?
This also includes software that you have used for any manual steps.
If the code has particular hardware requirements
(e.g., in terms of processor or memory),
these should be also noted.
Finally, for steps that take more than a couple of seconds,
the approximate runtime should be indicated.

@nte-dependencies provides more information
on determining the dependencies of an R project.

### Usage

How can one run the project -- is there a master script
or a particular order in which any scripts need to be executed?
Provide detailed instructions for running the full project.

### List of Results

For every result (i.e., number, figure, or table) that is
computed in the project and displayed in the manuscript,
indicate where exactly it is computed.

### Citation

Is there a recommended way to cite this project?
Is there a published article associated with it
that you would like to have cited?

### License

Under which licenses are the works in this project folder available?

## Create It!

Create your README now as the file `README.md`.

::: {#nte-dependencies .callout-note}
### Identifying R Dependencies

R itself and the R packages are already documented as this project uses `renv`.
Therefore you can focus on all other dependencies,
such as the system dependencies of R packages
as well as the version of Quarto.[^renv-quarto]

[^renv-quarto]: As of August 2024, a proposal for `renv` to record the version of Quarto
has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).

An overview over the system dependencies of R packages can be created
using the function `pak::pkg_sysreqs()`.
In combination with `renv`, we can obtain the system dependencies
of all R packages the current project directly depends on:

```{.r filename="Console"}
# Find all R package dependencies
deps <- renv::dependencies()$Package |>
  unique() |>
  pak::pkg_deps(dependencies = NA) |>
  getElement("package")

# Identify their system dependencies
pak::pkg_sysreqs(deps)
```

The output may look like the following:

```txt
── Install scripts ────────────────── Fedora 40 ── 
dnf install -y make pandoc git

── Packages and their system dependencies ────────
fs        – make
knitr     – pandoc
remotes   – git
rmarkdown – pandoc
sass      – make
```

We can see that the programs `make`, `pandoc`,
and `git` were identified as system dependencies.
Often, one can obtain their version by running them with the `--version` argument:

```{.bash filename="Terminal"}
make --version
pandoc --version
git --version
```

However, this does not work for all system dependencies.
Specifically, it does not work for libraries -- software that is not supposed to be run on its own.
Identifying their version is beyond the scope of this tutorial.

We also know that we need Quarto to create the PDF,
so let's find out its version as well:

```{.bash filename="Terminal"}
quarto --version
```

If you installed `apaquarto` or any other Quarto extension,
one can query their versions as follows:[^already-included]

[^already-included]: Luckily, the extensions are included in the project folder,
so technically their version is already recorded in the project's files.

```{.bash filename="Terminal"}
quarto list extensions
```

Finally, we know that we installed a $\TeX$ distribution to create the PDF,
so let's find out its version by running:

```{.bash filename="Terminal"}
quarto check
```

The output is quite long and it might look slightly different for you,
but the relevant sections are the following:

```txt
[✓] Checking tools....................OK
      TinyTeX: v2024.09
      Chromium: (not installed)

[✓] Checking LaTeX....................OK
      Using: TinyTex
      Path: /home/r155953/.TinyTeX/bin/x86_64-linux
      Version: 2024
```

Of course, all the system dependencies identified until now
may have dependencies on their own.
Use your own judgement to decide when not to dig deeper.
:::

@@

The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF accoording to APA guidelines.
```
:::

::: {#tip-data .callout-tip collapse="true"}
### Involved Data

```{.md .code-overflow-wrap filename="README.md"}
## Involved Data

@@

Next, one can open a new R session in the root folder of this project and run the following, which should install all required R packages at their recorded versions:

@@

::: {#tip-usage .callout-tip collapse="true"}
### Usage

`````{.md .code-overflow-wrap filename="README.md"}
## Usage

@@

::: {#tip-list-of-results .callout-tip collapse="true"}
### List of Results

```{.md .code-overflow-wrap filename="README.md"}
## List of Results

- In-text numbers in the section "results": Calculated in the chunk "t-test" within "Manuscript.qmd"
- Table 1: Calculated in the chunk "tbl-descriptive-statistics" within "Manuscript.qmd"
- Figure 2: Calculated in the chunk "fig-bill-length-comparison" within "Manuscript.qmd"
```
:::

@@

The manuscript files `Manuscript.qmd`, `Manuscript.tex`, and `Manuscript.pdf` by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) or (at your option) under the [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later). For further copyright information, see `LICENSE.txt`.
```
:::

>>>>

Having settled on a license, it is time to add a final touch.
Imagine returning to your project in five years,
having forgotten most of the details of what you did exactly.
What would your future self want to know in order to quickly understand
what is going on in the project?

This is what needs to be described in the `README` file.
It should be the primary entry point into your project,
the place where new users go first to get an orientation.

There is no common standard how to structure a README file,
but you should usually provide at least the following information,
structured by sections:

**Name and Description**

- [ ] How is the project called?
- [ ] What is it about?
- [ ] Which files does the project folder contain?
- [ ] How are they organized?

**Data**

- [ ] Are any (empirical) data involved (e.g., being analyzed or used as input)?
- [ ] Are they already included in the project folder? If not: From which sources can they be obtained?
- [ ] Where is their data dictionary located?
- [ ] Which terms, usage restrictions, or licenses apply?
- [ ] (optionally) If they are not publicly available, is an alternative, synthesized version provided?

**Computational Requirements / Dependencies**

- [ ] What software needs to be installed to run the analysis?

For every dependency (e.g., packages that you load in R), describe where it can be obtained from.
If the code has particular hardware requirements
(e.g., in terms of processor or memory), these should be also noted.
Finally, for steps that take more than a couple of seconds,
indicate the approximate runtime.

For an R project, the most important information for enabling reproducibility are the specific version of R and the loaded packages. Ideally, you also list other system dependencies. The in-depth supplementary material [*Identify additional system dependencies for the README*](/in_depth_material/other_dependencies.qmd) provides more information on determining the dependencies of an R project.
*For the purpose of this tutorial, however, you can skip this step.*

**How to reproduce the results**

- [ ] Provide step-by-step instructions that are necessary to reproduce all reported results.

Is there a master script or a particular order in which any scripts need to be executed?

**Citation**

- [ ] Is there a recommended way to cite this project?
- [ ] Is there a published article associated with it that you would like to have cited?

**License**

- [ ] Under which licenses are the works in this project folder available?

## Identifying R Dependencies

If you just write "We used R to analyze the data" in your manuscript or in the *Computational Requirements / Dependencies* section of your README file, this would be too underspecified. To enable someone else to really reproduce your results, you need specific information on ...

- The exact version of R
- All packages and their versions that you loaded in your project
- (optionally) Further system dependencies

### R & Package Dependencies

R itself and the R packages are already documented as our tutorial project uses `renv`: The information is stored in the `renv.lock` file at the root of the project, which tracks each package’s version and source. In this case, you can simply refer to that in the README file and all is done:

> The R and packages versions used to compute the results are stored in the `renv` environment (see `renv.lock` file).

If you do *not* use `renv`, the easiest solution to get a list of all needed packages is the `sessionInfo()` command. *Call it **after** you loaded all necessary packages!*

It lists all packages including their versions, and you can copy & paste that information into your README file. Here's an example for a project:

```txt
> sessionInfo()
R version 4.4.1 (2024-06-14)
Platform: aarch64-apple-darwin20
Running under: macOS 15.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: Europe/Berlin
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
 [1] rio_1.2.2            doRNG_1.8.6          rngtools_1.5.2      
 [4] doParallel_1.0.17    iterators_1.0.14     foreach_1.5.2       
 [7] fitPCurve_0.0.0.9000 testthat_3.2.3       prettycode_1.1.0    
[10] colorDF_0.1.7

loaded via a namespace (and not attached):
 [1] gtable_0.3.6       ggplot2_3.5.2      htmlwidgets_1.6.4  devtools_2.4.5    
 [5] remotes_2.5.0      TruncExpFam_1.2.1  vctrs_0.6.5        tools_4.4.1       
 [9] generics_0.1.4     tibble_3.2.1       pkgconfig_2.0.3    R.oo_1.26.0       
[13] data.table_1.17.4  RColorBrewer_1.1-3 desc_1.4.3         lifecycle_1.0.4 
```

## ✍️ Practical Exercise: Create your README!

Create your README now as the file `README.md`, located at the top level folder.

@@

The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF according to APA guidelines.
```
:::

::: {#tip-data .callout-tip collapse="true"}
### Data

```{.md .code-overflow-wrap filename="README.md"}
## Data

@@

Next, one can open a new R session in the root folder of this project and run the following command, which should install all required R packages at their recorded versions:

@@

::: {#tip-how-to-reproduce .callout-tip collapse="true"}
### How to reproduce the results

`````{.md .code-overflow-wrap filename="README.md"}
## How to reproduce the results

@@

@@

The manuscript files `Manuscript.qmd`, `Manuscript.tex`, and `Manuscript.pdf` by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/). For further copyright information, see `LICENSE.txt`.
```
:::
